---
title: "Reproducible Research: Peer Assessment 1"
author: "Marek Mañko"
date: "Sunday, March 15, 2015"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r echo=TRUE}
require(ggplot2)
strPath <- "E:\\Marek\\Studia\\Coursera\\Reproducible Research\\PA_1\\"
Imp.Activity <- read.csv(paste(strPath,"activity.csv", sep=""))
```

## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day
2. Make a histogram
3. Calculate and report the mean and median of the total number of steps taken per day
```{r echo=TRUE}
Proc.Activity <- aggregate(steps ~ date, Imp.Activity, sum, na.rm=TRUE)
g <- ggplot(Proc.Activity, aes(x=steps)) + geom_histogram(binwidth=500)
print(g)

# Mean
mean(Proc.Activity$steps)
# Median
median(Proc.Activity$steps)
```

## What is the average daily activity pattern?
1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r echo=TRUE}
Proc.Activity <- aggregate(steps ~ interval, Imp.Activity, mean, na.rm=TRUE)
g <- ggplot(Proc.Activity, aes(x=interval, y=steps)) + geom_line() + xlab("Time interval") + ylab("Mean Number of steps")
print(g)

# 5-min interval
Proc.Activity.MaxInterval <- Proc.Activity$interval[which.max(Proc.Activity$steps)]
```
5-minute interval which contains the maximum number of steps is interval 835.

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset.
2. Devise a strategy for filling in all of the missing values in the dataset.
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Comment on results.
```{r echo=TRUE}
# There are 2304 missing values for 'steps'
summary(Imp.Activity)
# Fill NAs with mean values for each interval calculated above
names(Proc.Activity)[2] <- "mean_steps"
Proc.NA_fill <- merge(Imp.Activity, Proc.Activity, by=c("interval"), all.x=TRUE)
Proc.NA_fill$steps <- ifelse(is.na(Proc.NA_fill$steps), Proc.NA_fill$mean_steps, Proc.NA_fill$steps)

# New dataset equal to the 
Imp.Activity.New <- Proc.NA_fill[,c("steps","date","interval")]
summary(Imp.Activity.New)

Proc.Activity.New <- aggregate(steps ~ date, Imp.Activity.New, sum, na.rm=TRUE)
g <- ggplot(Proc.Activity.New, aes(x=steps)) + geom_histogram(binwidth=500)
print(g)

# Mean
mean(Proc.Activity.New$steps, na.rm=TRUE)
# Median
median(Proc.Activity.New$steps, na.rm=TRUE)
# Conclusion: Daily mean remained almost unchanged, median is currently equal to mean.
```

## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend”
2. Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r echo=TRUE}
require(plyr)
Imp.Activity.New$date <- as.Date(Imp.Activity.New$date)
Imp.Activity.New$DayType <- as.factor(ifelse(weekdays(Imp.Activity.New$date) %in% c("sobota","niedziela"),"weekend", "weekday"))
Proc.Activity.New <- ddply(Imp.Activity.New, .(interval,DayType),summarize, steps = mean(steps))
g <- ggplot(Proc.Activity.New, aes(x=interval, y=steps)) + geom_line() + xlab("Time interval") + 
    ylab("Mean Number of steps") + facet_grid(. ~ DayType)
print(g)

```

